---
title: 'STAT 420: Final Project Report'
authors: "Team"
output:
  html_document:
    theme: readable
    toc: yes
  pdf_document:
    toc: yes
---


```{r}
#install.packages("mltools")
library(mltools)
library(data.table)
```

**Loading dataset in R from the CSV file, and converting **
```{r}
car_data = read.csv("Cars_data.csv")
car_data = subset(car_data, select = -c(Vehicle.Style, Market.Category))

#unique(car_data$Transmission.Type)
#colnames(car_data)
#unique(car_data$Engine.Fuel.Type)
#unique(car_data$Driven_Wheels)
unique(car_data$Vehicle.Size)
#head(car_data)

```

**Removing the dew non automatic/manual transmission types, and storing this new data in car_data_transd dataframe**
```{r}


car_data_transd<-car_data[!(car_data$Transmission.Type=="AUTOMATED_MANUAL" | car_data$Transmission.Type=="DIRECT_DRIVE" | car_data$Transmission.Type=="UNKNOWN"),]

unique(car_data_transd$Transmission.Type)

```


**Removing certain fuel types, keeping only gasoline and diesel. Storing the result in car_data_fuel dataframe**

```{r}
car_data_fuel<-car_data_transd[!(grepl("flex", car_data_transd$Engine.Fuel.Type, fixed = TRUE)
|car_data_transd$Engine.Fuel.Type=="electric" | car_data_transd$Engine.Fuel.Type=="" | car_data_transd$Engine.Fuel.Type=="natural gas"),]

unique(car_data_fuel$Engine.Fuel.Type)
```

**Assigning the different types of gasoline to a single "gasoline value". Now, the only two values for fuel type will be "gasoline" and "diesel" as visible below**

```{r}
car_data_fuel$Engine.Fuel.Type[car_data_fuel$Engine.Fuel.Type == "premium unleaded (required)" ] <- "gasoline"
car_data_fuel$Engine.Fuel.Type[car_data_fuel$Engine.Fuel.Type == "regular unleaded" ] <- "gasoline"
car_data_fuel$Engine.Fuel.Type[car_data_fuel$Engine.Fuel.Type == "premium unleaded (recommended)" ] <- "gasoline"

unique(car_data_fuel$Engine.Fuel.Type)

```



**One hot encoding - converting categorical variables to values which R will understand**
```{r}
car_data_factored = car_data_fuel
car_data_factored$Vehicle.Size <- factor(car_data_factored$Vehicle.Size)
car_data_factored$Transmission.Type <- factor(car_data_factored$Transmission.Type)
car_data_factored$Engine.Fuel.Type <- factor(car_data_factored$Engine.Fuel.Type)
car_data_factored$Driven_Wheels <- factor(car_data_factored$Driven_Wheels)

levels(car_data_factored$Vehicle.Size)
levels(car_data_factored$Transmission.Type)
levels(car_data_factored$Engine.Fuel.Type)
levels(car_data_factored$Driven_Wheels)



car_data_factored = one_hot(as.data.table(car_data_factored))

```



**Added a variable to dataset (ReleasedYearsAgo) which stores the value of the years ago this model was released**

```{r}
car_data_factored$ReleasedYearsAgo <- with(car_data_factored, 2020 - Year)
```



```{r}
head(car_data_factored)
```

